### 2014-05-08 paper/content.tex
changed things to paper

\documentclass[11pt,a4paper]{article}
%\documentclass[conference]{IEEEtran}
\usepackage[english]{babel} 
\usepackage[T1]{fontenc}
\usepackage{pdfpages}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage{eurosym}
\DeclareUnicodeCharacter{20AC}{\euro}
\usepackage{csquotes}
\usepackage{graphicx}
\usepackage{rotating} 
\usepackage{array}
\usepackage{booktabs} 
\usepackage{longtable}
\usepackage{ifthen}
\usepackage{xcolor}
\usepackage{tabu}
\usepackage{colortbl}
\usepackage{calc}
\usepackage{pifont}
\usepackage{forloop}
\usepackage[nomessages]{fp}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\newtheoremstyle{break}
   {\topsep}{\topsep}%
   {\itshape}{}%
   {\bfseries}{}%
   {\newline}
   {\thmname{#1}\thmnumber{\@ifnotempty{#1}{ }\@upn{#2}}%
    \thmnote{ {\bfseries(#3)}}}%

\newtheorem{definition}{Definition}

\usepackage[parfill]{parskip}
%\usepackage[onehalfspacing]{setspace}
\usepackage{newclude}

\newcounter{starnumber}
\newcommand{\stars}[1]{
  \forloop{starnumber}{1}{\value{starnumber} < 6}{
    \ifthenelse{#1 < \value{starnumber}}{\ding{73}}{\ding{72}}%
  }
}

\renewcommand{\arraystretch}{1.2}

%\usepackage[sortcites=true, style=authoryear,natbib=true]{biblatex}
\usepackage{biblatex}
\bibliography{bibliography}
\nocite{*}

%Autornamen in Bibliography fett
%\AtBeginBibliography{\renewcommand*{\mkbibnamelast }[1]{\textbf{#1}}}
%\AtBeginBibliography{\renewcommand*{\mkbibnamefirst }[1]{\textbf{#1}}}

%\makeindex

\title{Development of a educational and sustainable robotic platform for children and adults}
\date{\today}
\author{Sebastian Muszytowski \\Baden-Wuerttemberg Cooperative State University Mannheim }

\begin{document}

\maketitle
%\twocolumn[
%       \maketitle
%        \begin{@twocolumnfalse}
%        \maketitle
%        \end{@twocolumnfalse}
%]

\begin{abstract}
Kinder, Jugendliche und Erwachsene an Technik heranzuführen ist eine wichtige Aufgabe in einer zunehmend technisierten Gesellschaft. Ein verbreiteter Ansatz ist die Heranführung mit Hilfe von Robotern da diese ihre Umgebung wahrnehmen und mit dieser interagieren können.

Eine Evaluation zeigt, dass viele Robotikplattformen für den Bildungssektor Kriterien an eine nachhaltige und pädagogisch wertvolle Lösung zu Teilen bereits erfüllen. Kriterien wie günstige Anschaffungskosten, Erweiterbarkeit und geringe Komplexität sind nur zu Teilen erfüllt und zeigen die Notwendigkeit einer neuen Plattform.

Nach einer detalierten Analyse zu Bewegungskonzepten, Sensoren, Kommunikation, Mikrocontroller-Plattformen und Batterie-Systemen wird anhand einer Auswahl von Komponenten eine kostengünstige, pädagogisch wertvolle, quelloffene und erweiterbare Robotikplattform konstruiert. Eine detalierte Darstellung der Entwicklungsschritte zeigt auf, welche Designentscheidung die Bewertungskriterien positiv beeinflussen.

Die Abschließende Evaluation von Chancen und Risiken zeigt, dass eine pädagogisch sinnvolle und wirtschaftliche Verwendung der neu entwickelten Robotikplattform möglich ist. Die zuvor auf andere Plattformen angewendeten Evaluationskriterien zeigen eine deutliche Verbesserung in Punkten Erweiterbarkeit, Anschaffungskosten und Nachhaltigkeit.
\end{abstract}

\begin{abstract}
Getting children, young adults and adults in touch with technology is important with today’s technologically advancing society. A common approach is to use Robots since they can sense their environment and modify it.

The evaluation of different robotic platforms designed for educational use shows that criteria such as sustainability and educational value are met to a certain extend. Other criteria such as low cost of purchase, extendibility and low complexity are partially met and therefore show the need for a new robotic platform.

After a detailed analysis of movement concepts, sensors, communication techniques, microcontroller platforms and power supply systems the optimal components for the new  platforms were chosen. The choice is based on cost, educational, extension and open source criteria. A detailed description of the construction steps allows to comprehend design decisions which positively influence the evaluation criteria.

A evaluation of chances and challenges on the constructed robot shows that a educational use of the platform is possible whilst operating economically. The previously used evaluation criteria for evaluating other platforms applied to the new platform shows enhancements in extendibility, purchase costs and sustainability.
\end{abstract}

\section{Introduction}
Getting children in touch with technology is important in todays technologically advanced society. In schools technology (specifically electrical engineering and programming) is often taught using robots. Robots are well suited for children since they allow interaction with the reality in contrast to arbitrary technology related exercises such as programming exercises which doesn't affect the real world.\newline 
Since schools are required to save money, robots used for teaching stay property of the schools. Therefore children usually cannot take the robot home to work with it. In addition modification of robot is usually not permitted which reduces potential learning experiences. Giving students the full control over all possibilities of the robot requires them to own the robot. This requires to robot to be cheap, extensible and suitable for educational purposes.\newline
The development of a robot platform for educational purpose requires requirement engineering, market research and evaluation of robots which are already used in education. Once the research is completed a new robot platform can be constructed taking the research results into account. The goal is to develop such an educational robot platform with regard to the target user group, their needs and educational concepts.
\section{Requirement Analysis / Project Goals}
User group of a robot can be categorized into different groups regardless of their age:
\begin{description}
\item[Beginner] \hfill \\ Beginners have little to no experience with robots. They need extensive assistance to perform tasks such as soldering and programming. For programming graphical programming languages (such as minibloq\footnote{\url{http://blog.minibloq.org/}} or ModKit\footnote{\url{http://www.modkit.com/}}) or languages providing a high level of abstraction (such as Wiring), 
\item[Advanced] \hfill \\ Advanced users have advanced experience with robots and robotics. Most programming and soldering tasks are done independently; working knowledge is utilized to solve new and unknown tasks.  
\item[Expert] \hfill \\ Expert users understand all aspects of robots and their internals. New challenges are faced with ease.
\end{description}

@@

In the following chapters the state of the art in robotics is evaluated, robotic components are assessed and the construction of the robot according to the requirements is presented in detail and critically discussed. The structure is aimed to give a big picture and get into detail as more and more components are introduced.
\section{State of the Art}
The development of a new robot platform requires the assessment of current robotic platforms using the requirements developed. In focus of the assessment are popular robot platforms which are either actively used in schools or developed with education in mind. All of the platforms are independently developed and focus on different ideas.

@@

\subsection{Asuro}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/asuro.jpg}
  \caption{Picture of the Asuro (CC-BY-SA Robin Gruber)}
\end{figure}

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{5}    & Having a price tag of 50 Euro renders the Asuro very affordable in comparison to the 100 Euro limit set. The cost effectiveness is outstanding and features sensors and actuators for basic tasks.     \\
Educational     & \stars{3}     & Asuro features detailed build-instructions in English and German. The guide also includes a tutorial on how to program Asuro in C. Even though there are a lot of opportunities to teach electrical engineering, the available documents do not explain the electrical usage and functional principle of each part. \\
Sustainable       & \stars{4}     & Asuro uses standard components only and features ROHS compliant parts which are free from heavy metals such as lead (PB). A broken Asuro can be easily repaired using the tools which are required to solder it. \\
Extendable & \stars{2}      & Having no free microcontroller pins and no broken out pins, Asuro cannot be considered extendable. There exist some hardware modification in the Asuro community which extend the functionality but require to unsolder/remove some parts.  \\
Trouble-free & \stars{3} & Building Asuro is well documented but some steps may lead to frustration. Mounting the axis using solder is very difficult and may burn the PCB due to the heat emitted in this process. Another frustrating task is to program the robot using the infrared serial connection, since this method is prone to interference. \\
Open-Source & \stars{4} & Asuro was licensed under the DLR-license which is transitioned into a Creative-Commons license model. Everything except the PCB layout files are available for download on the Asuro website.\\
\bottomrule
\caption{Asuro evaluation}
\label{tbl:asuro_eval}
\end{longtable}

@@

\subsection{Arduino Robot}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/arduinorobot.jpg}
  \caption{Picture of the Arduino Robot}
\end{figure}

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{2}    & With a price tag of 189 Euro without VAT\footnote{As of 05.03.2014 from the official Arduino store: \url{http://store.arduino.cc/index.php?main_page=product_info&cPath=11_12&products_id=290}} the Arduino Robot is far above the price limit of 100 Euro. The price for the basic robot is very high compared to its features.\\
Educational     & \stars{3}     & The Arduino robot comes pre-built and therefore lacks the build experience. Learning material for teachers is not provided but a lot of examples and a library to ease programming is provided. Instructions for other Arduino products can be easily adapted to the robot. \\
Sustainable       & \stars{4}     &  The robot integrates charging circuit for the NIMH batteries eliminates the need of a dedicated battery charger and therefore reduces the amount of electrical waste. A small restraint is the choice of components because they cannot be interchanged/repaired by beginners. This is due to the manufacturing process which requires SMD components to be cost effective. Through hole parts would be easier to interchange but require manual assembly which is very expensive.\\
Extendable & \stars{5}      &  Having prototyping areas directly on the control board, the robot can be considered extendable. Moreover connectors for additional sensors are populated all over the robot board making it possible and easy to add sensors without soldering.\\
Trouble-free & \stars{4} & Since the Arduino robot comes pre-assembled and tested in factory, the robot can be called trouble-free. Transportation defects like loose screws can be repaired using household items. The usage is very simple and straight-forward.\\
Open-Source & \stars{3} & Even though it is claimed that the Arduino robot is open-source, not all files required for building the robot are published. The website is missing the PCB design files and includes the schematics only. A software library is available under a open-source license. \\
\bottomrule
\caption{Arduino Robot evaluation}
\label{tbl:arduinorobot_eval}
\end{longtable}

@@

\subsection{LEGO Mindstorms NXT 2.0}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/mindstorms.jpg}
  \caption{Picture of the LEGO Mindstorms NXT (from http://www.devoxx.com/)}
\end{figure}

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{2}    & With a price tag of 279 Euro (retail consumer edition) and 368.99 Euro (education version) the robot is far beyond the price tag of 100 Euro. The amount of sensors and actuators is low compared to the price.\\
Educational & \stars{5}     & LEGO provides a lot of resources for teaching and personal education using the Mindstorms NXT 2.0. Even though the electrical components are not subject, the set can teach a lot of knowledge and even interdisciplinary content. The flexibility of the robot allows the transformation of the robert for a lot of education concern. \\
Sustainable  & \stars{3}     & The Mindstorms NXT 2.0 set cannot be repaired using household items. Nevertheless the components are considered sustainable to a certain extend, since they are designed to be robust and to last a long time.\\
Extendable & \stars{3} & The robot can be extended up to a certain extend. Even though the specification for each interface is public available custom sensors are rare and/or expensive. \\
Trouble-free & \stars{5} & Using the platform is trouble free and easy. The LEGO Mindstorms IDE is intuitional. Support for different programming languages such as Java exist. A huge community and a lot of resources are available.\\
Open-Source & \stars{4} & LEGO open sourced the LEGO Mindstorms NXT brick and moreover the specification for each connector and sensor. The PCB design files and the casing are not open-sourced to prevent unauthorized replicas.\\
\bottomrule
\caption{Lego Mindstorms evaluation}
\label{tbl:mindstorms_eval}
\end{longtable}

@@

\subsection{Thymio II}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/thymioii.jpg}
  \caption{Picture of the Thymio II (from https://aseba.wikidot.com/en:thymio)}
\end{figure}

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{4}    & 99 CHF (without taxes directly from the Ecole Polytechnique Fédérale de Lausanne) is a reasonable price for a robot with this huge amount of features. On retail stores the price is higher due to profit margin.\\
Educational & \stars{3}     & Thymio II does not yet have a repository of educational material only, but it is planned to release such material. The compability with LEGO is a huge bonus, and creates a lot of pplications on which educational problems can be explained and demonstrated.\\
Sustainable  & \stars{3}     & The robot consists of small SMD parts and therefore parts cannot be replaced by the target user group. Materials used are suitable for children and designed to last. The integrated Li-Po battery can be recharged and therefore produces less environmental waste.\\
Extendable & \stars{3} & Except the LEGO compatible fixation points, Thymio II is not extendable without unsoldering components. On the software side, the choice of microcontroller supports a lot of additional program features.\\
Trouble-free & \stars{4} & Thymio II is trouble free since it comes preassembled and the usage is straight forward.\\
Open-Source & \stars{5} & Everything needed to build the robot is licensed under a permissive creative commons license. Program files and libraries are licensed under a LGPL license. Even the CAD files for the plastic mold injection parts are available without a charge.\\
\bottomrule
\caption{Thymio II evaluation}
\label{tbl:thymio_eval}
\end{longtable}

@@

\section{Case studies}
A robot is built using parts and modules which can be grouped into functional groups like movement, sensors or communication. The goal of each functional group can be achieved in various ways which have to be evaluated using the requirements used in the evaluation of other robot platforms. Some evaluation critera such as the fulfillment of the educational aspect or the open-source aspect may not be suitable for evaluation of single components and therefore left out. The Extendability describes the amount of microcontroller pins where the lesser amount of pins is considered better since this leaves space for extensions. Other difficulties in the evaluation process are dependencies between different modules, e.g. when a sensor has special power requirements. Based on the evaluation the robot will be constructed.
\subsection{Movement}
Movement is the most important feature since it allows the robot to interact with its environment. In contrast to a stationary robot, a moving robot can interact with a wider area but requires battery power or similar mobile energy sources. Different approaches for robot movement exist which will be part of critical evaluation.

@@

Typically robots have two wheels and a ball caster which prevents the robot from falling over. A ball caster acts as a third unidirected wheel with reduced friction. Alternatively four wheels and a steering mechanism or three wheels using omniwheels can be used. All types of wheeled robots can move in flat terrain and tolerate slight uneven surfaces. The outdoor usage is limited.
\subsubsection{Wheeled with two DC motors}
DC motors (direct current motors) are a cheap solutions for driving wheels. Usually a gear box is required to reduce the motor speed in favor of torque. Controlling the direction (forward and backward) requires the ability to switch the direction of the current flow. This is usally archieved using a H-bridge which can be found in as an integrated circuit or build using four electrical switches.

@@

\subsubsection{Wheeled with two Stepper motors}
Stepper motors are basically DC motors with more coils. In stepper motors the full rotation is divided into a number of equal steps which can be separately controlled. The motor can step forward, backward or hold the current position with huge torque. Stepper motors are usually shipped within a gear box which reduces steps much further to gain higher precision and torque. 
In typical applications stepper motors are driven by a dual H-bridge or darlington arrays attached to shift registers. The latter method is very cost effective but limited in current supply, the first method utilized a motor driver such as the L293 per stepper.

@@

If a stepper motor gets broken, it can be easily replaced using the same model of stepper motor. Since the wheel is not part of the stepper motor, the old wheel can be replaced easily.
\subsubsection{Omniwheels}
Omniwheels/Mecanum wheels are the most flexible way of moving using wheels. Omniwheels are a special form of wheels which allow the movement in any direction having three independently moveable wheels. It requires three motors (typically DC motors) with direction control and optional speed control.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/mecanum.png}
  \caption{Screenshot of a mecanum wheel 3D model (from http://www.thingiverse.com/thing:2318)}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/omniwheel.jpg}
  \caption{Control scheme of a omni wheel (from http://commons.wikimedia.org/wiki/File:Omni\_wheel\_control.jpg)}
  \label{fig:omnicontrol}
\end{figure}

\subsubsection{Conclusion}
All in all a solution using stepper motors with darlington array connected to shift registers is the most cost-effective solution. It provides the necessary functions and complies to the evaluation aspects such as the educational potential. Its usage is rather trouble free when apropriate libraries for motor control are provided.

\subsection{Sensors}
Sensors are important to make the robot sense its environment. Sensing the environment enables interaction such as obstacle avoidance, finding objects or following lines. Connecting sensors to a microcontroller can be done in various ways using various protocols. Different protocols require a different amount of microcontroller pins. Dependening on the sensor technology more or less computing time is needed to evaluate the sensors value.

\subsubsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever has an integrated Schmitt trigger and therefore provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

@@

\subsubsection{IR Proximity Sensor}
Infrared proximity sensors can be used in various ways. When measuring distances using proximity sensors one has to choose the right sensor type since the range varies from a few centimeters to a few meters. Another use of IR proximity sensors is to measure brightness. This usually works with sensors designed for very small distances. When the IR light is emitted it reflects better on bright surfaces than on dark surfaces which do not reflect light. A photodiode for infrared wavelength changes its electrical behaviour on the influence of the light reflected which can be measured using the microcontrollers analog-to-digital converter.

@@

\subsubsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29m/s at sea level).

@@

\subsubsection{Accelerometer}
Accelerometers can sense acceleration on one or more axis. This can be used to determine the speed of the robot or the direction in which the robot accelerates (which can be used to determine whether the robot falls/is lifted up).

@@

The MPU-6050 can be interfaced using various protocols, but I2C is the most common way to do so and requires two input/output pins only using software-based I2C or two I2C hardware pins. Libraries with permissive license exist which ease the use of the accelerometer. For advanced usage scenarios a integrated DMP (Digitial Motion Processor) can process raw values and interface additional sensors like a compass.
\subsubsection{Gyroscope}
Gyroscopes are used to determine the orientation. Applications for gyroscopes are balancing and position calculation (usually in combination with an accelerometer). Integrated circuits such as the MPU-6050 mentioned before usually combine a gyroscope and an accelerometer into one single device.

@@

As mentioned in the previous section the MPU-6050 is easily interfaced using various protocols of which I2C is the common way.
\subsubsection{Compass}
A compass allows the robot to sense its orientation relative to the magnetic field of the earth. Technically speaking a integrated circuit providing measurement of the magnetic field of the earth are a special form of magnetometers.

@@

\subsubsection{RFM12B}
The RFM12B is a wireless transceiver operating in the 434MHz or 915MHz band. It is produced by HOPE RF and provides an SPI interface to communicate with a microcontroller.

@@

\subsection{Electrical requirement}
Each component has its own electrical requirement such as voltage, current and pin count. The requirements for all components can be added up to create a power supply and a wiring concept.

>>>>

\chapter{Introduction}
Getting children in touch with technology is important in todays technologically advanced society. In schools technology (specifically electrical engineering and programming) is often taught using robots. Robots are well suited for children since they allow interaction with the reality in contrast to arbitrary technology related exercises such as programming exercises which doesn't affect the real world.\cite[pp.16]{johnson2003children}\newline 
Since schools are required to save money, robots used for teaching stay property of the schools. Therefore children usually cannot take the robot home to work with it. In addition modification of robot is usually not permitted which reduces potential learning experiences. Giving students the full control over all possibilities of the robot requires them to own the robot. This requires to robot to be cheap, extensible and suitable for educational purposes.\newline
The development of a robot platform for educational purpose requires requirement engineering, market research and evaluation of robots which are already used in education. Once the research is completed a new robot platform can be constructed taking the research results into account. The goal is to develop such an educational robot platform with regard to the target user group, their needs and educational concepts.
\chapter{Requirement Analysis / Project Goals}
User group of a robot can be categorized into different groups regardless of their age:
\begin{description}
\item[Beginner] \hfill \\ Beginners have little to no experience with robots. They need extensive assistance to perform tasks such as soldering and programming. For programming graphical programming languages (such as minibloq\footnote{\url{http://blog.minibloq.org/}} or ModKit\footnote{\url{http://www.modkit.com/}}) or languages providing a high level of abstraction (such as Wiring), 
\item[Advanced] \hfill \\ Advanced users have advanced experience with robots and robotics. Most programming and soldering tasks are done independently; working knowledge is utilized to solve new and unknown tasks.  
\item[Expert] \hfill \\ Expert users understand all aspects of robots and their internals. New challenges are faced with ease.
\end{description}

@@

In the following chapters the state of the art in robotics is evaluated, robotic components are assessed and the construction of the robot according to the requirements is presented in detail and critically discussed. The structure is aimed to give a big picture and get into detail as more and more components are introduced.
\chapter{State of the Art}
The development of a new robot platform requires the assessment of current robotic platforms using the requirements developed. In focus of the assessment are popular robot platforms which are either actively used in schools or developed with education in mind. All of the platforms are independently developed and focus on different ideas.

@@

\section{Asuro}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/asuro.jpg}
  \caption{Picture of the Asuro (CC-BY-SA Robin Gruber)}
\end{figure}

@@

\begin{longtabu} to \textwidth { X[1,l] X[1,l] X[4,l]}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{5}    & Having a price tag of 50 Euro renders the Asuro very affordable in comparison to the 100 Euro limit set. The cost effectiveness is outstanding and features sensors and actuators for basic tasks.     \\
Educational     & \stars{3}     & Asuro features detailed build-instructions in English and German. The guide also includes a tutorial on how to program Asuro in C. Even though there are a lot of opportunities to teach electrical engineering, the available documents do not explain the electrical usage and functional principle of each part. \\
Sustainable       & \stars{4}     & Asuro uses standard components only and features ROHS compliant parts which are free from heavy metals such as lead (PB). A broken Asuro can be easily repaired using the tools which are required to solder it. \\
Extendable & \stars{2}      & Having no free microcontroller pins and no broken out pins, Asuro cannot be considered extendable. There exist some hardware modification in the Asuro community which extend the functionality but require to unsolder/remove some parts.  \\
Trouble-free & \stars{3} & Building Asuro is well documented but some steps may lead to frustration. Mounting the axis using solder is very difficult and may burn the PCB due to the heat emitted in this process. Another frustrating task is to program the robot using the infrared serial connection, since this method is prone to interference. \\
Open-Source & \stars{4} & Asuro was licensed under the DLR-license which is transitioned into a Creative-Commons license model. Everything except the PCB layout files are available for download on the Asuro website.\\
\bottomrule
\caption{Asuro evaluation}
\label{tbl:asuro_eval}
\end{longtabu}

@@

\section{Arduino Robot}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/arduinorobot.jpg}
  \caption{Rendering of the Arduino Robot (CC-BY-SA arduino.cc)}
\end{figure}

@@

\begin{longtabu} to \textwidth { X[1,l] X[1,l] X[4,l]}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{2}    & With a price tag of 189 Euro without VAT\footnote{As of 05.03.2014 from the official Arduino store: \url{http://store.arduino.cc/index.php?main_page=product_info&cPath=11_12&products_id=290}} the Arduino Robot is far above the price limit of 100 Euro. The price for the basic robot is very high compared to its features.\\
Educational     & \stars{3}     & The Arduino robot comes pre-built and therefore lacks the build experience. Learning material for teachers is not provided but a lot of examples and a library to ease programming is provided. Instructions for other Arduino products can be easily adapted to the robot. \\
Sustainable       & \stars{4}     &  The robot integrates charging circuit for the NIMH batteries eliminates the need of a dedicated battery charger and therefore reduces the amount of electrical waste. A small restraint is the choice of components because they cannot be interchanged/repaired by beginners. This is due to the manufacturing process which requires SMD components to be cost effective. Through hole parts would be easier to interchange but require manual assembly which is very expensive.\\
Extendable & \stars{5}      &  Having prototyping areas directly on the control board, the robot can be considered extendable. Moreover connectors for additional sensors are populated all over the robot board making it possible and easy to add sensors without soldering.\\
Trouble-free & \stars{4} & Since the Arduino robot comes pre-assembled and tested in factory, the robot can be called trouble-free. Transportation defects like loose screws can be repaired using household items. The usage is very simple and straight-forward.\\
Open-Source & \stars{3} & Even though it is claimed that the Arduino robot is open-source, not all files required for building the robot are published. The website is missing the PCB design files and includes the schematics only. A software library is available under a open-source license. \\
\bottomrule
\caption{Arduino Robot evaluation}
\label{tbl:arduinorobot_eval}
\end{longtabu}

@@

\section{LEGO Mindstorms NXT 2.0}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/mindstorms.jpg}
  \caption{Picture of the LEGO Mindstorms NXT (CC-BY-NC Elvis Ripley)}
\end{figure}

@@

\begin{longtabu} to \textwidth { X[1,l] X[1,l] X[4,l]}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{2}    & With a price tag of 279 Euro (retail consumer edition) and 368.99 Euro (education version) the robot is far beyond the price tag of 100 Euro. The amount of sensors and actuators is low compared to the price.\\
Educational & \stars{5}     & LEGO provides a lot of resources for teaching and personal education using the Mindstorms NXT 2.0. Even though the electrical components are not subject, the set can teach a lot of knowledge and even interdisciplinary content. The flexibility of the robot allows the transformation of the robert for a lot of education concern. \\
Sustainable  & \stars{3}     & The Mindstorms NXT 2.0 set cannot be repaired using household items. Nevertheless the components are considered sustainable to a certain extend, since they are designed to be robust and to last a long time.\\
Extendable & \stars{3} & The robot can be extended up to a certain extend. Even though the specification for each interface is public available custom sensors are rare and/or expensive. \\
Trouble-free & \stars{5} & Using the platform is trouble free and easy. The LEGO Mindstorms IDE is intuitional. Support for different programming languages such as Java exist. A huge community and a lot of resources are available.\\
Open-Source & \stars{4} & LEGO open sourced the LEGO Mindstorms NXT brick and moreover the specification for each connector and sensor. The PCB design files and the casing are not open-sourced to prevent unauthorized replicas.\\
\bottomrule
\caption{Lego Mindstorms evaluation}
\label{tbl:mindstorms_eval}
\end{longtabu}

@@

\section{Thymio II}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/thymioii.jpg}
  \caption{Picture of the Thymio II (CC-BY-SA Thymio II project team)}
\end{figure}

@@

\begin{longtabu} to \textwidth { X[1,l] X[1,l] X[4,l]}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{4}    & 99 CHF (without taxes directly from the Ecole Polytechnique Fédérale de Lausanne) is a reasonable price for a robot with this huge amount of features. On retail stores the price is higher due to profit margin.\\
Educational & \stars{3}     & Thymio II does not yet have a repository of educational material only, but it is planned to release such material. The compability with LEGO is a huge bonus, and creates a lot of pplications on which educational problems can be explained and demonstrated.\\
Sustainable  & \stars{3}     & The robot consists of small SMD parts and therefore parts cannot be replaced by the target user group. Materials used are suitable for children and designed to last. The integrated Li-Po battery can be recharged and therefore produces less environmental waste.\\
Extendable & \stars{3} & Except the LEGO compatible fixation points, Thymio II is not extendable without unsoldering components. On the software side, the choice of microcontroller supports a lot of additional program features.\\
Trouble-free & \stars{4} & Thymio II is trouble free since it comes preassembled and the usage is straight forward.\\
Open-Source & \stars{5} & Everything needed to build the robot is licensed under a permissive creative commons license. Program files and libraries are licensed under a LGPL license. Even the CAD files for the plastic mold injection parts are available without a charge.\\
\bottomrule
\caption{Thymio II evaluation}
\label{tbl:thymio_eval}
\end{longtabu}

@@

\chapter{Case studies}
A robot is built using parts and modules which can be grouped into functional groups like movement, sensors or communication. The goal of each functional group can be achieved in various ways which have to be evaluated using the requirements used in the evaluation of other robot platforms. Some evaluation critera such as the fulfillment of the educational aspect or the open-source aspect may not be suitable for evaluation of single components and therefore left out. The Extendability describes the amount of microcontroller pins where the lesser amount of pins is considered better since this leaves space for extensions. Other difficulties in the evaluation process are dependencies between different modules, e.g. when a sensor has special power requirements. Based on the evaluation the robot will be constructed.
\section{Movement}
Movement is the most important feature since it allows the robot to interact with its environment. In contrast to a stationary robot, a moving robot can interact with a wider area but requires battery power or similar mobile energy sources. Different approaches for robot movement exist which will be part of critical evaluation.

@@

Typically robots have two wheels and a ball caster which prevents the robot from falling over. A ball caster acts as a third unidirected wheel with reduced friction. Alternatively four wheels and a steering mechanism or three wheels using omniwheels can be used. All types of wheeled robots can move in flat terrain and tolerate slight uneven surfaces. The outdoor usage is limited.
\subsection{Wheeled with two DC motors}
DC motors (direct current motors) are a cheap solutions for driving wheels. Usually a gear box is required to reduce the motor speed in favor of torque. Controlling the direction (forward and backward) requires the ability to switch the direction of the current flow. This is usally archieved using a H-bridge which can be found in as an integrated circuit or build using four electrical switches.

@@

\subsection{Wheeled with two Stepper motors}
Stepper motors are basically DC motors with more coils. In stepper motors the full rotation is divided into a number of equal steps which can be separately controlled. The motor can step forward, backward or hold the current position with huge torque. Stepper motors are usually shipped within a gear box which reduces steps much further to gain higher precision and torque. 
In typical applications stepper motors are driven by a dual H-bridge or darlington arrays attached to shift registers. The latter method is very cost effective but limited in current supply, the first method utilized a motor driver such as the L293 per stepper.

@@

If a stepper motor gets broken, it can be easily replaced using the same model of stepper motor. Since the wheel is not part of the stepper motor, the old wheel can be replaced easily.
\subsection{Omniwheels}
Omniwheels/Mecanum wheels are the most flexible way of moving using wheels. Omniwheels are a special form of wheels which allow the movement in any direction having three independently moveable wheels. It requires three motors (typically DC motors) with direction control and optional speed control.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/mecanum.png}
  \caption{Screenshot of a mecanum wheel 3D model (from \url{http://www.thingiverse.com/thing:2318})}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/omniwheel.jpg}
  \caption{Control scheme of a omni wheel (from \url{http://commons.wikimedia.org/wiki/File:Omni\_wheel\_control.jpg})}
  \label{fig:omnicontrol}
\end{figure}

\subsection{Conclusion}
All in all a solution using stepper motors with darlington array connected to shift registers is the most cost-effective solution. It provides the necessary functions and complies to the evaluation aspects such as the educational potential. Its usage is rather trouble free when apropriate libraries for motor control are provided.

\section{Sensors}
Sensors are important to make the robot sense its environment. Sensing the environment enables interaction such as obstacle avoidance, finding objects or following lines. Connecting sensors to a microcontroller can be done in various ways using various protocols. Different protocols require a different amount of microcontroller pins. Dependening on the sensor technology more or less computing time is needed to evaluate the sensors value.

\subsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever has an integrated Schmitt trigger and therefore provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

@@

\subsection{IR Proximity Sensor}
Infrared proximity sensors can be used in various ways. When measuring distances using proximity sensors one has to choose the right sensor type since the range varies from a few centimeters to a few meters. Another use of IR proximity sensors is to measure brightness. This usually works with sensors designed for very small distances. When the IR light is emitted it reflects better on bright surfaces than on dark surfaces which do not reflect light. A photodiode for infrared wavelength changes its electrical behaviour on the influence of the light reflected which can be measured using the microcontrollers analog-to-digital converter.

@@

\subsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29m/s at sea level).

@@

\subsection{Accelerometer}
Accelerometers can sense acceleration on one or more axis. This can be used to determine the speed of the robot or the direction in which the robot accelerates (which can be used to determine whether the robot falls/is lifted up).

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_gyroaccel.jpg}
  \caption{MPU-6050 breakout board (CC-BY-SA Ben Oswald)}
\end{figure}

The MPU-6050 can be interfaced using various protocols, but I2C is the most common way to do so and requires two input/output pins only using software-based I2C or two I2C hardware pins. Libraries with permissive license exist which ease the use of the accelerometer. For advanced usage scenarios a integrated DMP (Digitial Motion Processor) can process raw values and interface additional sensors like a compass.
\subsection{Gyroscope}
Gyroscopes are used to determine the orientation. Applications for gyroscopes are balancing and position calculation (usually in combination with an accelerometer). Integrated circuits such as the MPU-6050 mentioned before usually combine a gyroscope and an accelerometer into one single device.

@@

As mentioned in the previous section the MPU-6050 is easily interfaced using various protocols of which I2C is the common way.
\subsection{Compass}
A compass allows the robot to sense its orientation relative to the magnetic field of the earth. Technically speaking a integrated circuit providing measurement of the magnetic field of the earth are a special form of magnetometers.

@@

\subsection{RFM12B}
The RFM12B is a wireless transceiver operating in the 434MHz or 915MHz band. It is produced by HOPE RF and provides an SPI interface to communicate with a microcontroller.

@@

\section{Electrical requirement}
Each component has its own electrical requirement such as voltage, current and pin count. The requirements for all components can be added up to create a power supply and a wiring concept.
